Deep Learning

### 2015-01-07 deepreview.tex
added language paragraph and working on cleaning references

From the earliest days of pattern recognition~\cite{selfridge, rosenblatt}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
  Rumelhart, schmidreference}.  Backpropagation is simply an application of
the chain rule, which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights in a layered
network. The objective function typically measures the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it outputs} of a layer can be computed by working
backwards from the error derivatives with respect to the outputs of the
subsequent layer. Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer (see box
1).

@@

The hidden layers of a multi-layer neural network learn to represent the
network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}. 
Each context word is presented to the network as a one-of-N
vector in which one component has a value of $1$ and the rest are $0$. The
output layer has one unit per word and is forced to produce a probability
distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden
layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
\end{equation}

@@

The family tree demonstration used a small, clean dataset in which the
regularities could be captured by micro-rules that were never violated and
the features were easy to interpret, but learned word-vectors also work
very well when the word sequences come from a large corpus of real text and
the individual micro-rules are unreliable\cite{BenDucVin01-short}. When trained to
predict the next word in a news story, for example, the learned word
vectors for Monday and Tuesday are very similar, as are the word vectors
for Vietnam and Iraq.  Vector representations of words learned from text
are now very widely used in natural language applications\cite{Collobert,
  Mikolov, Socher?}.

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}.

@@

After your stuff on language, I would like to include some version of the
paragraph below along with a figure.
%%%%%%%%%%%%%%%%%%%%%%%

When translating between languages, the hidden units of the decoder network
that generates possible translations are initalized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see figure ??).

@@

\subsection{Back-propagation and Stochastic Gradient Descent}
Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
533-536.

>>>>

From the earliest days of pattern recognition~\cite{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s\cite{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  Backpropagation is simply an application of
the chain rule, which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights in a layered
network. The objective function typically measures the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it outputs} of a layer can be computed by working
backwards from the error derivatives with respect to the outputs of the
subsequent layer. Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer (see box
1).

@@

The hidden layers of a multi-layer neural network learn to represent the
network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}. 
Each context word is presented to the network as a one-of-N
vector in which one component has a value of $1$ and the rest are $0$. The
output layer has one unit per word and is forced to produce a probability
distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{\sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden
layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ \exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
\end{equation}

@@

The family tree demonstration used a small, clean dataset in which the
regularities could be captured by micro-rules that were never violated and
the features were easy to interpret, but learned word-vectors also work
very well when the word sequences come from a large corpus of real text and
the individual micro-rules are unreliable~\citep{BenDucVin01-short}. When trained to
predict the next word in a news story, for example, the learned word
vectors for Monday and Tuesday are very similar, as are the word vectors
for Vietnam and Iraq.  Vector representations of words learned from text
are now very widely used in natural language applications~\cite{Collobert,
  Mikolov, Socher?}.

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. The elements of such vectors can be understood as
the values of learned attributes (or features) associated with some object. 
With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}.

@@

After your stuff on language, I would like to include some version of the
paragraph below along with a figure.
%%%%%%%%%%%%%%%%%%%%%%%
\fi

Prior to the introduction of neural language
models~\citet{BenDucVin01-short} the standard approach to statistical
modeling of language was based on counting frequencies of occurences of
short symbol sequences (called n-grams). One disadvantage is that it
taking into account context beyond a handful of words would require
training corpora of exponentially growing size (with respect to the
desired context size), since the number of configurations of $n$
words from a vocabulary of size $V$ grows as $V^n$. Related to this
problem and the very notion of symbol introduced above, n-grams
do not generalize across the many semantically related sequences 
of words (for example the many ways of expressing roughly the same thing).
Instead, neural language models learn to associate each word to a
distributed representation (a vector) such that semantically related
words end up close to each other in that space, as illustrated in
Figure~\ref{fig:word-embeddings}. To obtain that effect, one
{\em jointly learns} a representation for each word as well as
a function that computes a quantity of interest that is relevant
to a language related task, such as predicting the next word
in a sequence (neural language models) or predicting the sequence
of translated words given the sequence of source words, i.e., neural
machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
\caption{Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}.}
\label{fig:word-embeddings}
\end{figure}

When translating between languages, the hidden units of the decoder network
that generates possible translations are initalized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see Figure~\ref{fig:caption-generation}).

\begin{figure}[ht]
%\centerline{\includegraphics[width=0.7\linewidth]{}}
\caption{Example of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from an image, with the whole system trained to ``translate''
images into captions. Reproduced from~\citet{?}.
}
\label{fig:caption-generation}
\end{figure}

@@

\subsection{Back-propagation and Stochastic Gradient Descent}

%Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
%representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
%533-536.
\citep{Rumelhart86b}
